---
title: "Catch-at-age"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: FALSE
    code_folding: show
    number_sections: TRUE
---
    
# SET-UP
```{r  markdown, echo=F}
library(knitr)
opts_chunk$set(echo = T, collapse = T, fig.align = "center", fig.width = 9, fig.height = 6)
options(width = 120)
```    

## settings
```{r settings, message = F}
source('0.0_settings.R')
```

## load all data
```{r data, message = F}
load(paste0(dir.rdat, "catch.Rdata"))                                # from 1.0 read
load(paste0(dir.rdat, "bio.Rdata"))                                  # from 1.0 read
load(paste0(dir.rdat, "lf.Rdata"))                                   # from 1.0 read
group.region <- read.table(paste0(dir.dat,'caa_group_region.txt'),header=T)
group.gear <- read.table(paste0(dir.dat,'caa_group_gear.txt'),header=T)
```

# LENGTH-WEIGHT RELATIONSHIPS 

Sample weights are often missing and apparently are not precise. Therefore, the weight of a sample is estimated based on the length of the individuals and their number sampled. This section fits length-weight relationships for all years and periods. Bio samples are used ('carbio').

```{r lw, message = F}

## subset
lw <- bio[!is.na(bio$length) & !is.na(bio$weight),c('year','trim','length','weight')]
lw$logw <- log(lw$weight)           
lw$logl <- log(lw$length)          

## remove extremes
# ggplot(lw,aes(x=logl,y=logw))+geom_point()
lw <- lw[lw$logw>0 & lw$logl>3,]                         # minimal weight and length 

## relationships by year and period (quarter) (!robuust regression instead of removing subjectively some points each year!)
lw.modR <- paste0(dir.rdat, "lw.mod.Rdata")

if(!file.exists(lw.modR)){
    lw.mod <- with(lw,fit.lw(logl,logw,year,trim,tresh=300,period.unit='quarter'))
    save(lw.mod, file =  lw.modR)
}else{ 
    load(lw.modR)
}
```

## CHECK {.tabset}
### Prediction
```{r lw fit}
check <- expand.grid.df(lw.mod,data.frame(length=0:ceiling(max(lf$length))))
check$weight <- with(check,exp(intercept+slope*log(length))*cf)
ggplot(check,aes(x=length,y=weight,col=year,group=year))+
    geom_line(size=0.5)+
    facet_wrap(~period)+
    scale_y_continuous(limits=c(0,max(lw$weight)))+
    scale_color_viridis_c()+
    labs(title='Predicted length-weight relationships')
```


# LENGTH-FREQUENCY DATA

```{r lf, message = F, fig.height = 3,fig.width = 4}
## subset
lf.caa <- lf[,c('year','trim','nafo','gear.cat','length','n','sample.id')]

## group regions and gears
# PERIOD: quarter
names(lf.caa)[2] <- 'period'

# REGION: large scale regions 
lf.caa$region <- group.region[match(lf.caa$nafo,group.region$nafo),'region']
lf.caa$nafo <- NULL

# GEAR: similar selectivity gears
lf.caa$gear <- group.gear[match(lf.caa$gear.cat,group.gear$gear.cat),'gear.group']
lf.caa$gear.cat <- NULL

## convert lengths to sample weights (kg)
lf.caa <- merge(lf.caa,lw.mod,all.x=T)
lf.caa$weight.sample <- with(lf.caa,exp(intercept+slope*log(length))*cf*n/100)

## make equal length classes
table(lf$length.bin)
lf.caa$length <-roundFish(lf.caa$length,5)  

save(lf.caa, file =  paste0(dir.rdat, "lf.caa.Rdata"))

```

## CHECK {.tabset}

### model samples
```{r table lw, echo=FALSE}
check <- unique(lf.caa[,c('year','period','option')])
kable(with(check,table(option,period)))
```
### predictions
```{r plot lw, echo=FALSE, fig.height = 3,fig.width = 10}
grid.arrange(
  ggplot(lf.caa,aes(x=weight.sample/n))+geom_histogram(binwidth = 1)+labs(title='Predicted unit weights',x='Weight (kg)'),
  ggplot(lf.caa,aes(x=weight.sample))+geom_histogram(binwidth = 10)+labs(title='Predicted sample weights',x='Weight (kg)'),
  ncol=2
)
```

## TABLES {.tabset}
```{r lf tab1}
df <- lf.caa
df$id <- with(df,paste(sample.id,year,period,region,gear),collapse='.')
```
### period
```{r lf tab2}
dfp <- ddply(df,c('year','period'),summarise,N=length(unique(id)),n=sum(n))
dfp <- dcast(dfp,year~period,value.var='N',fill = 0)
dfp$total <- rowSums(dfp[,-1])
kable(dfp)
```
### region
```{r lf tab3}
dfr <- ddply(df,c('year','region'),summarise,N=length(unique(id)),n=sum(n))
dfr <- dcast(dfr,year~region,value.var='N',fill = 0)
dfr$total <- rowSums(dfr[,-1])
kable(dfr)
```
### gear
```{r lf tab4}
dfg <- ddply(df,c('year','gear'),summarise,N=length(unique(id)),n=sum(n))
dfg <- dcast(dfg,year~gear,value.var='N',fill = 0)
dfg$total <- rowSums(dfg[,-1])
kable(dfg)
```

# BIO DATA

```{r bio, message = F, fig.height = 3,fig.width = 4}
## subset
bio.caa <- bio[!is.na(bio$agef) & !is.na(bio$length),
               c('year','trim','nafo','gear','length','agef','sample.id')]
names(bio.caa)[6] <- 'age'

table(bio.caa$age)
bio.caa <- bio.caa[bio.caa$age<20,]        # remove crazy ages (42 and 63)

## group regions and gears
# PERIOD: quarter
names(bio.caa)[2] <- 'period'
table(bio.caa$period,useNA = 'always')

# REGION: large scale regions 
bio.caa$region <- group.region[match(bio.caa$nafo,group.region$nafo),'region']
bio.caa$nafo <- NULL
table(bio.caa$region,useNA = 'always')

# GEAR: similar selectivity gears
bio.caa$gear <- group.gear[match(bio.caa$gear,group.gear$gear.cat),'gear.group']
table(bio.caa$gear,useNA = 'always')

## make equal length classes
bio.caa$length <- roundFish(bio.caa$length,5) 
bio.caa <- bio.caa[bio.caa$length>0,]
range(bio.caa$length)

save(bio.caa, file =  paste0(dir.rdat, "bio.caa.Rdata"))
```

## TABLES {.tabset}
```{r bio tab1, echo=F}
df <- bio.caa
df$id <- with(df,paste(sample.id,year,period,region,gear),collapse='.')
```
### period
```{r bio tab2, echo=F}
dfp <- ddply(df,c('year','period'),summarise,N=length(unique(id)),n=sum(n))
dfp <- dcast(dfp,year~period,value.var='N',fill = 0)
dfp$total <- rowSums(dfp[,-1])
kable(dfp)
```
### region
```{r bio tab3, message = F}
dfr <- ddply(df,c('year','region'),summarise,N=length(unique(id)),n=sum(n))
dfr <- dcast(dfr,year~region,value.var='N',fill = 0)
dfr$total <- rowSums(dfr[,-1])
kable(dfr)
```
### gear
```{r bio tab4, message = F}
dfg <- ddply(df,c('year','gear'),summarise,N=length(unique(id)),n=sum(n))
dfg <- dcast(dfg,year~gear,value.var='N',fill = 0)
dfg$total <- rowSums(dfg[,-1])
kable(dfg)
```

# CATCH DATA

```{r catch.caa, message = F, fig.height = 8,fig.width = 6}
catch <- catch[,c('year','month','nafo','gear.cat','source','catch')]

## identify period, region and gear
# 1) PERIOD: quarters
catch$period <- ceiling(catch$month/3)

# 2) REGION: large scale regions 
catch$region <- group.region[match(catch$nafo,group.region$nafo),'region']

# 3) GEAR: similar selectivity gears
catch$gear <- group.gear[match(catch$gear.cat,group.gear$gear.cat),'gear.group']

## total catch by level
catch.caa <- ddply(catch,c('year','region','period','gear'),summarise,catch=sum(catch,na.rm = T))
catch.caa <- catch.caa[catch.caa$catch>0,]

save(catch.caa, file =  paste0(dir.rdat, "catch.caa.Rdata"))

# plot
p.r <- plotCatch(catch.caa,x='year',y='catch',fill='region')
p.g <- plotCatch(catch.caa,x='year',y='catch',fill='gear')
p.p <- plotCatch(catch.caa,x='year',y='catch',fill='period')
grid.arrange(p.r,p.g,p.p)
```

# CATCH-AT-AGE

```{r caa, message = F}
caa.allR <- paste0(dir.rdat, "caa.all.Rdata")
caaR <- paste0(dir.rdat, "caa.Rdata")
  
if(!all(file.exists(c(caa.allR,caaR)))){
    # step 1) get samples
    lf.caa[,which(names(lf.caa) %in% names(lw.mod)[-c(1:2)])] <- NULL       # remove model details
    catch.caa <- catch.caa[catch.caa$year %in% unique(lf.caa$year),]        # years for which bio/lf
    
    caa.all <- caa.raw(catch=catch.caa,
                       lf=lf.caa, 
                       al=bio.caa,
                       tresh.lf=2,          # min samples for length-frequency
                       tresh.al=2,          # min samples for age-length key
                       period.unit='quarter')   # quarterly grouping instead of monthly

    caa.all$X0 <- NULL                       # remove age 0 ?? (total landings somewhat smaller)
    save(caa.all, file =  caa.allR)
    
    # step 2) get actual caa
    caa <- caa.clean(caa.all,plus=10)
    save(caa, file =  caaR)     
}else{ 
    load(caa.allR)
    load(caaR)
}
```

## CHECK {.tabset}
### samples
```{r caa check samples, fig.height = 6,fig.width = 8}
check <- ddply(caa.all,c('id'),summarise,
               option.lengthfreq=unique(option.lengthfreq),
               nsample.lengthfreq=unique(nsample.lengthfreq),
               nfish.lengthfreq=sum(n.lf),
               option.agelength=unique(option.agelength),
               nsample.agelength=unique(nsample.agelength),
               nfish.agekey=sum(n.agekey))
check <- melt(check,id='id')
ggplot(check,aes(x=value))+geom_histogram(bins=30)+facet_wrap(~variable,scale='free')  # region and gear can only rarely be matched
```
### total
```{r caa check total, fig.height = 6,fig.width = 8}
tot <- ddply(catch.caa,c('year'),summarise,catch=sum(catch))
ggplot(caa,aes(x=year))+
  geom_bar(stat='identity',aes(y=caaw,fill=age))+
  geom_line(data=tot,aes(y=catch))+
  scale_fill_viridis()
```

# PLOTS {.tabset}

## CAA

```{r caa plot, message = F, fig.height = 6,fig.width = 16}
grid.arrange(
  ggplot(caa,aes(x=year,y=age,size=caan))+geom_point(alpha=0.8)+ scale_size(range = c(1,10)),
  ggplot(caa,aes(x=year,y=age,size=caaw))+geom_point(alpha=0.8)+ scale_size(range = c(1,10)),
  ncol=2
)
```

## WAA

```{r waa plot, message = F, fig.height = 6,fig.width = 16}
grid.arrange(
  ggplot(caa,aes(x=year,y=waa,group=age))+
    geom_line(aes(color=age))+
    scale_color_viridis(),
  ggplot(caa,aes(x=year,y=waa,group=age))+
    geom_ribbon(aes(ymin=waa-2*waa.sd,ymax=waa+2*waa.sd,fill=age),alpha=0.2)+
    geom_line(aes(color=age))+
    scale_color_viridis(),
  ncol=2
)

```
